---
title: "ST_512-homework2"
author: "Lindsay Dodson"
format: html
---

## Problem 1

|            |     |       |        |         |         |
|------------|-----|-------|--------|---------|---------|
| source     | df  | SS    | MS     | F-value | p-value |
| dosage/trt | 2   | 450.7 | 225.35 | 7.031   | 0.95    |
| error      | 9   | 288.2 | 32.02  |         |         |
| total      | 11  | 738.9 | 257.4  |         |         |

: Lindsay Dodson

```{r}
# f-value
f1<- qf(1-.95, 2, 11)
f1
1-(pf(f1,2,9))
```
**a**

Because F>f, we reject the null hypothesis, and have evidence to conclude that dosage level does affect bioactivity.

**b**

```{r}
t_val1<-qt(.975,11)
point_est1<- 36.75
SE1<- 1.634

point_est1 - t_val1*SE1
point_est1 + t_val1*SE1
```

The 95% confidence interval is (33.153, 40.346). 

**c**
```{r}
t_val2<-qt(.995, 11)
point_est2<- 44.75-29.75
SE2<- sqrt((32.02)*((1/4)+(1/4)))

point_est2 - t_val2*SE2
point_est2 + t_val2*SE2
```

The 99% confidence interval is (2.57, 27.42).

**d**

simultaneous CI ??

**e**

```{r}
#Standard Error 
SE3<-(32.02/4)*2

#Point Estimate (from part c)
point_est2

#Tukeys
tuk_1<-(qtukey(1-0.05, 3, 9))/sqrt(2) #critical value
tuk_1*SE3 #margin of error

point_est2+(tuk_1*SE3) #Confidence interval
point_est2-(tuk_1*SE3)

#Bonferoni
bon_1<-qt(1-(0.05/6),9) #CV
bon_1*SE3 #MoE

point_est2+(bon_1*SE3) #CI
point_est2-(bon_1*SE3)

#Schaffe's
schaf_1<- sqrt((qf(1-0.05,2,9))*2) #CV
schaf_1*SE3 #MoE

point_est2+(schaf_1*SE3) #CI
point_est2-(schaf_1*SE3)

```
Confidence Intervals

Tukey: (-29.7,59.7)

Bonferoni: (-31.96, 61.96)

Schaffe's: (-31.71, 61.71)

**e**

Tukey's HSD = $Q*\sqrt{\frac{MSE}{n}}$ 

HSD=5.26

1 vs 3: $|29.75-44.75|=15 > 5.26$ different

1 vs 2: $|29.75-36.75|=7 > 5.26$ different

2 vs 3: $|36.75-44.75|=8 > 5.26$ different


## Problem 2

```{r}
strength<- c(3129, 3000, 2865, 2890, 3200, 3300, 2975, 3150, 2800, 2900, 2985, 3050, 2600, 2700, 2600, 2765) 
mix_tech<- factor(rep(c("1", "2", "3", "4"), each = 4))
Portland_data <- data.frame(mix_tech, strength)
print(Portland_data)

anova_model <- aov(strength ~ mix_tech, data = Portland_data)
summary(anova_model)
```
**a**

Because the p-value=0.0005<<alpha=0.05, we reject the null hypothesis and have strong evidence to conclude that the mixing techniques affect the strength of the cement. 

**b**




