---
title: "ST_512-homework2"
author: "Lindsay Dodson"
format: html
---

## Problem 1

|            |     |       |        |         |         |
|------------|-----|-------|--------|---------|---------|
| source     | df  | SS    | MS     | F-value | p-value |
| dosage/trt | 2   | 450.7 | 225.35 | 7.031   | 0.95    |
| error      | 9   | 288.2 | 32.02  |         |         |
| total      | 11  | 738.9 | 257.4  |         |         |

: Lindsay Dodson

```{r}
# f-value
f1<- qf(1-.95, 2, 11)
f1
1-(pf(f1,2,9))
```

**a**

Because F\>f, we reject the null hypothesis, and have evidence to conclude that dosage level does affect bioactivity.

**b**

```{r}
t_val1<-qt(.975,11)
point_est1<- 36.75
SE1<- 1.634

point_est1 - t_val1*SE1
point_est1 + t_val1*SE1
```

The 95% confidence interval is (33.153, 40.346).

**c**

```{r}
t_val2<-qt(.995, 11)
point_est2<- 44.75-29.75
SE2<- sqrt((32.02)*((1/4)+(1/4)))

point_est2 - t_val2*SE2
point_est2 + t_val2*SE2
```

The 99% confidence interval is (2.57, 27.42).

**d**

```{r}
#Standard Error 
SE3<-(32.02/4)*2

#Point Estimate (from part c)
point_est2

#Tukeys
tuk_1<-(qtukey(1-0.05, 3, 9))/sqrt(2) #critical value
tuk_1*SE3 #margin of error

point_est2+(tuk_1*SE3) #Confidence interval
point_est2-(tuk_1*SE3)

#Bonferroni
bon_1<-qt(1-(0.05/6),9) #CV
bon_1*SE3 #MoE

point_est2+(bon_1*SE3) #CI
point_est2-(bon_1*SE3)

#Scheffe's
schaf_1<- sqrt((qf(1-0.05,2,9))*2) #CV
schaf_1*SE3 #MoE

point_est2+(schaf_1*SE3) #CI
point_est2-(schaf_1*SE3)

```

Confidence Intervals

Tukey: (-29.7,59.7)

Bonferroni: (-31.96, 61.96)

Scheffe's: (-31.71, 61.71)

**e**

Tukey's HSD = $Q*\sqrt{\frac{MSE}{n}}$

HSD=5.26

1 vs 3: $|29.75-44.75|=15 > 5.26$ different

1 vs 2: $|29.75-36.75|=7 > 5.26$ different

2 vs 3: $|36.75-44.75|=8 > 5.26$ different

## Problem 2

```{r}
strength<- c(3129, 3000, 2865, 2890, 3200, 3300, 2975, 3150, 2800, 2900, 2985, 3050, 2600, 2700, 2600, 2765) 
mix_tech<- factor(rep(c("1", "2", "3", "4"), each = 4))
Portland_data <- data.frame(mix_tech, strength)
print(Portland_data)

anova_model <- aov(strength ~ mix_tech, data = Portland_data)
summary(anova_model)
```

**a**

Because the p-value=0.0005\<\<alpha=0.05, we reject the null hypothesis and have strong evidence to conclude that the mixing techniques affect the strength of the cement.

**b**

```{r}
mean_mix3<- 2933.75
mix_CV<- qt(1-.05, 15) ##is alpha 0.05 or .95
SE_mix<- sqrt(12825.7/16) # check SE

mean_mix3 - (mix_CV*SE_mix)
mean_mix3 + (mix_CV*SE_mix)
```

The confidence interval is (2884.12, 2983.38)

**c**

```{r}
## techniques 1 & 3
mean_mix1 <- 2971.00
PE_13<- (mean_mix1-mean_mix3)

PE_13 - (mix_CV*SE_mix)
PE_13 + (mix_CV*SE_mix)

## techniques 3 & 4
mean_mix4 <- 2666.25
PE_34<- (mean_mix3-mean_mix4)

PE_34 - (mix_CV*SE_mix)
PE_34 + (mix_CV*SE_mix)
```

The confidence interval for techniques 1 & 3 is (-12.39, 86.89).

The confidence interval for techniques 3 & 4 is (217.87, 317.13).

**d**

```{r}
#Tukey's
tuk_portland<-(qtukey(1-0.05, 4, 12))/sqrt(2) #can we use the same SE ? sd is the same for all groups, so that would make SE 0 if not using MSE/n forumla
tuk_portland*SE_mix
PE_13-(tuk_portland*SE_mix)
PE_13+(tuk_portland*SE_mix)

#Bonferroni
bon_portland<- qt((1-.0041667),12)
bon_portland*SE_mix

PE_13-(bon_portland*SE_mix)
PE_13+(bon_portland*SE_mix)

#Scheffe's
sch_portland<- sqrt(3*qf(1-0.05,3,12))
sch_portland*SE_mix

PE_13-(sch_portland*SE_mix)
PE_13+(sch_portland*SE_mix)
```

Tukey's CI: (-46.81, 121.31)

Bonferroni CI: (-52.01, 126.51)

Scheffe's CI: (-54.37, 128.87)

**e**

```{r}
bon_portland*SE_mix
```

**f**

```{r}
tuk_portland*SE_mix
```

**g**

```{r}
sch_portland*SE_mix
```

## Problem 3

Contrast 1:

$L_1=(1)\mu_1+(-\frac{1}{3})\mu_2+(-\frac{1}{3})\mu_3+(-\frac{1}{3})\mu_4=0$

$\hat{L}=\bar{y}_1\cdot-\frac{\bar{y}_2\cdot+\bar{y}_3\cdot+\bar{y}_4\cdot}{3}=1.867$

$sd(\hat{L_2})=MSE*\sum\frac{c_i^2}{n_i} = 196.0*0.067=13.07$

Contrast 2:

$L_1=(1)\mu_2+(-\frac{1}{2})\mu_3+(-\frac{1}{2})\mu_4=0$

$\hat{L}=\bar{y}_2\cdot-\frac{\bar{y}_3\cdot+\bar{y}_4\cdot}{2}=1.6$

$sd(\hat{L_2})=MSE*\sum\frac{c_i^2}{n_i} = 196.0*0.075=14.7$

Contrast 3:

$L_1=(1)\mu_3+(-1)\mu_4=0$

$\hat{L}=\bar{y}_3\cdot-\bar{y}_4\cdot=1.6$

$sd(\hat{L_2})=MSE*\sum\frac{c_i^2}{n_i} = 196.0*0.1=19.6$

## Problem 4
```{r}
radon_release<-c(80,83,83,85,75,75,79,79,74,73,76,77,67,72,74,74,62,62,67,69,60,61,64,66)
diameter<-c(rep("0.37",4),rep("0.51",4),rep("0.71",4),rep("1.02",4),rep("1.40",4),rep("1.99",4))
radon_data<-data.frame(radon_release,diameter)
radon_data
str(radon_data)

```

